P.berghei pir analysis

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
r functions moved to new folder

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

@@

@@

```{r function_uniquify}
#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
    name <- paste(vector_names[i],
                  df_num$rep[match(vector_names[i], df_num$names)],
                  sep = '_')
    #Increase rep counter
    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
    new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

@@

```{r function_ordering}
#Use an order_vector of partial string matches to order.
ordering = function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}

order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
                 'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')
r_order_vector <- str_replace_all(order_vector, 
                                  pattern = '-', 
                                  replacement = '\\.')
```

```{r tpm, results='hide'}

# We need to write a function to calculate the TPM from the count data we have generated. This function uses the _lengths.transcript_ object and should generate errors if it detects discrepancies between the Geneids so that the wrong transcript length is not used.

tpm = function(count_data,lengths) {

  #Print an error if the id columns do not match
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'id', 
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'id', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      stop('Warning:Count data and length gene ids DO NOT MATCH')
    }
  } else {
    stop('Count data and/or length table supplied do not have an id column, make sure they are ordered correctly')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'id',ignore.case = T)) == T){
    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'id', ignore.case = T)]
  } else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
  }else{
    lengths.table <- lengths}

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j,]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  if(any(grepl(colnames(count_data), pattern = 'id', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'id', ignore.case = T), drop = FALSE], 
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

>>>>

#Want the defaults for arrange, select, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

@@

```{r sourcing_functions}
source('../rfunctions/uniquify_fun.R')
source('../rfunctions/ordering_fun.R')
source('../rfunctions/tpm_fun.R')
```

@@

@@

```{r ordering_vectors}
order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
                 'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')

r_order_vector <- str_replace_all(order_vector, 
                                  pattern = '-', 
                                  replacement = '\\.')
```
